Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-02-17 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
finished section on event selection; completed sec iv.A for now

% \hs{Say that we use GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp} and justify why.}

%We will consider the events GW150915~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}.
%
%They are suitable for ringdown studies due to the high-masses and high signal-to-noise ratios (SNR).
The \pSEOB{} model, as described in Sec.~\ref{sec:review_pSEOB}, is an \emph{IMR} model which infers the properties of the underlying GW signal, including independently its ringdown properties, using the Bayesian formalism above. Naturally, the most promising candidates for our analyses are high-mass \emph{and} loud GW observations with a significant signal-to-noise ratio (SNR) in the post-merger signal. The latest LVK GW catalog~\cite{LIGOScientific:2021djp} reported 90 observed signals not all of which are relevant for a BH ringdown analysis. In fact, in an accompanying paper~\cite{LIGOScientific:2021sio}, the \textit{pSEOBNRv4HM} analysis~\footnote{Sec.VIII A.2 in ~\cite{LIGOScientific:2021sio}}, which is most similar to the \pSEOB{} model presented in this paper, identified two events which provided the strongest bounds on measurements of the dominant $(220)$ QNM -- GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}. These two events, with a total mass of $XX \Mo$ and $XX \Mo$ respectively, are extremely similar in their source properties. These are also two of the loudest BBH signals observed to date with a total network SNR of XX and XX respectively. Moreover, and what is most relevant for our analysis, is that their SNRs in the just the post-inspiral (merger-ringdown) part of the signal are the loudest to date, XX and XX respectively (corresponding tables in TGR GWTC-1/3) making them significantly promising as probes of the BH ringdown.

%Although the LIGO-Virgo detectors have observed 90 GW signals, we select, for our analyses in this paper, the ones that are likely to be the most promising probes of the BH ringdown using our method. In~\cite{LIGOScientific:2021sio}, the \textit{pSEOBNRv4HM} analysis in Sec.VIII A.2 (which is most similar to the \pSEOB{} model presented in this paper) identified two events which provided the strongest bounds on the dominant QNM measuments -- GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}.

\ag{event selection, noise, PE settings}

>>>>

The \pSEOB{} model, as described in Sec.~\ref{sec:review_pSEOB}, is an \emph{IMR} model which infers the properties of the underlying GW signal, including (independently) its ringdown properties, using the Bayesian formalism above. Naturally, the most promising candidates for our analyses are high-mass \emph{and} loud GW observations with a significant signal-to-noise ratio (SNR) in the post-merger part of the signal. The latest LVK GW catalog~\cite{LIGOScientific:2021djp} reported 90 observed signals not all of which are relevant for a BH ringdown analysis. In fact, in an accompanying paper~\cite{LIGOScientific:2021sio}, the \textit{pSEOBNRv4HM} analysis~\footnote{Sec.VIII A.2 in ~\cite{LIGOScientific:2021sio}}, which is most similar to the \pSEOB{} model presented in this paper, identified two events which provided the strongest bounds on measurements of the dominant $(220)$ QNM -- GW150914~\cite{LIGOScientific:2016aoc} and GW200129~\cite{LIGOScientific:2021djp}. These two events, with a total mass of $XX \Mo$ and $XX \Mo$ respectively, are extremely similar in their source properties. These are also two of the loudest BBH signals observed to date with a total network SNR of XX and XX respectively. Moreover, and what is more relevant for our analysis, is their post-inspiral (merger-ringdown) SNRs, which at XX and XX respectively (see the columns for $\rho_{\text{post-insp}}$ in Table III of~\cite{LIGOScientific:2019fpa} and Table IV of ~\cite{LIGOScientific:2021sio}). In this paper, we are going to focus on these two GW events as our probes of the BH ringdown in alternative theories of gravity.

The parameter inference in this paper follow configurations identical to the ones used on these events for the \textit{pSEOBNRv4HM} analysis in~\cite{LIGOScientific:2021sio}. GW150914 was a 2-detector event (HL) while GW200129 was 3-detector (HLV). We consequently use the same strain data $h(t)$, detector power-spectral-densities $S_n(f)$ and calibration envelopes as were used for the analyses in~\cite{LIGOScientific:2021sio}. The only difference between the sampling configurations is that we do not sample over the fractional deviations in the $(220)$ QNM frequencies and damping times, but instead on $\ell$ while holding $\{\delta \omega_k^{(j)}\},$ and $\{\delta \tau_k^{(j)}\}$ fixed to theory-specific predictions. In the section to follow, we enumerate through the different theories and outline the highlight results. Wherever possible, we also combine results from multiple events to obtain the strongest possible bound on $\ell$.
